\section{Related Work}
% Relevant work: distributed consensus algorithms (e.g., Paxos, Raft, Byzantine fault tolerance), UAV swarm resilience and cooperative autonomy, safety-critical systems, safety-critical dependability approaches in distributed systems.
This project will build on three main strands of existing research:
\begin{itemize}
    \item UAV swarms for search and rescue and related coverage missions.
    \item Distributed consensus and situation awareness in multi-agent systems.
    \item Dynamic task allocation and communication architectures for UAV networks.
\end{itemize}
The works reviewed in this section were selected because they either address SAR-oriented swarm behaviour, propose mechanisms for consensus or cooperative decision-making among UAVs, or survey task allocation and communication patterns relevant to resilient swarm coordination.

Together, these contributions provide the technical background for IRDS: They show how swarms can be organised, how information can be disseminated, and how tasks can be distributed. At the same time, they reveal important gaps with respect to health-aware, fail-operational replanning in SAR scenarios, where degraded agents must hand over responsibilities without compromising coverage or safety.

The most relevant characteristics of the reviewed works, including their mission focus, coordination strategy, and treatment of failures, are summarised in \cref{tab:placeholder}. These works were chosen because they collectively cover cooperative SAR missions, dynamic task allocation in swarms, consensus formation in UAV networks, and communication architectures for flying ad-hoc networks, which are all directly related to the IRDS problem of dependable replanning.
\begin{table}[H]
    \centering
    \begin{tabular}{|B{0.7cm}|C{5cm}|C{5cm}|C{5cm}|}
        \hline
        \textbf{Year} & \textbf{Title} & \textbf{Main Focus} & \textbf{Relevance to \acrshort{irds}} \\
        \hline
        2025 & Design of a Fail-Operational Swarm of Drones for Search and Rescue Missions \cite{LuizNotPublished} & High-level blueprint for a conceptual architecture for a fail-operational UAV swarm. & Degraded health message. \\
        \hline
        2023 & Cooperative Search and Rescue with Drone Swarm \cite{Luiz2023} & Distributed coordination and search resilience in UAV SAR missions. & Defines the application domain and justifies decentralised swarm cooperation for dependability. \\
        \hline
        2021 & Autonomous and Collective Intelligence for UAV Swarm in Target Search Scenario \cite{Luiz2021} & Conceptual architecture for autonomous \& collective intelligence in target search. & Serves as a theoretical base. \\
        \hline
        2025 & Development of Adaptive Drone Swarm Networks \cite{ChenYang-Yi2025DoAD} & Adaptive network layer (AeroSyn) & Suggests implementation model for health-state messaging \\ %Chen et al. (2025)
        \hline
        2025 & Dynamic reconnaissance operations with UAV swarms: adapting to environmental changes \cite{StodolaPetr2025Drow} & Dynamic replanning under failure & Provides comparative benchmark and validation model \\ %Stodola et al. (2025)
        \hline
        2025 & Energy Efficient Scheduling for Position Reconfiguration of Swarm Drones \cite{LiuHan2025EESf} & Energy-efficient reconfiguration & Supports energy-aware task redistribution \\ %Liu et al. (2025)
        \hline
        2023 & A Novel Distributed Situation Awareness Consensus Approach for UAV Swarm Systems \cite{HaiXingshuo2023ANDS} & Distributed consensus & Algorithmic basis for swarm agreement \\ %Hai et al. (2023)
        \hline
        2022 & A Review of Consensus-based Multi-agent UAV Implementations \cite{LizzioFaustoFrancesco2022ARoC} & Consensus-based UAV control & Practical implementation insights \\ %Lizzio et al. (2022)
        \hline
        2021 & Review of Dynamic Task Allocation Methods for UAV Swarms Oriented to Ground Targets \cite{QiangPeng2021RoDT} & Dynamic task allocation & Framework for reallocation strategy \\ %Peng et al. (2021)
        \hline
        2019 & UAV swarm communication and control architectures: a review \cite{CampionMitch2019Usca} & Swarm communication architectures & Justifies decentralised, \acrshort{fanet}-based design \\ %Campion et al. (2018)
        \hline
        %1 & 2 & 3 & 4 \\
        %\hline
    \end{tabular}
    \caption{Previous work used as refenced in the project.}
    \label{tab:placeholder}
\end{table}

The following subsections summarise the most relevant findings and limitations.

\subsection{Cooperative Search and Rescue with Drone Swarm}
This paper presents a cooperative search-and-rescue (SAR) strategy using UAV swarms.
It explores collective intelligence, distributed coordination, and fault-tolerant cooperation among \acrshortpl{uav} in dynamic SAR missions. The swarm shares situational data to maintain coverage and efficiency while adapting to \acrshort{uav} losses or degraded conditions.

\subsection{Autonomous and Collective Intelligence for UAV Swarm in Target Search Scenario}
This paper investigates collective intelligence architectures for UAV swarms performing target-search missions.
It introduces autonomous decision layers combining individual UAV autonomy (sensing, navigation) with collective coordination for search coverage.

\subsection{Development of Adaptive Drone Swarm Networks}
This paper introduces AeroSyn, a hybrid network architecture that combines cellular and \gls{adhoc} links. It enables \acrshortpl{uav}s to switch between connected and leader-follower communication modes to maintain stable swarm coordination.

\subsection{Dynamic reconnaissance operations with UAV swarms: adapting to environmental changes}
This paper develops a dynamic mission-replanning framework for UAV swarm reconnaissance using \acrfull{aco}. The model handles two events that force the system to adapt:
\begin{enumerate}
    \item Swarm composition changes (e.g., \acrshort{uav} failures/additions).
    \item Mission or environment changes (e.g., new area of responsibility).
\end{enumerate}

\subsection{Energy Efficient Scheduling for Position Reconfiguration of Swarm Drones}
This study proposes an energy-balancing reconfiguration scheme for \acrshort{uav} swarms operating in urban wind fields. The model determines when and how \acrshortpl{uav} should exchange positions to equalise energy use and extend swarm flight time.

\subsection{A Novel Distributed Situation Awareness Consensus Approach for UAV Swarm Systems}
This paper proposes a distributed situation awareness consensus framework that allows UAVs to share perception and align decision-making without a central controller. It introduces a dual-loop feedback mechanism to maintain robust cognitive agreement.

\subsection{A Review of Consensus-based Multi-agent UAV Implementations}
This review analyses consensus-based distributed control for \acrshort{uav} swarms. The review highlights practical issues in formation flight and target tracking, focusing on how communication delay and sensing accuracy affect practical use.

\subsection{Review of Dynamic Task Allocation Methods for UAV Swarms Oriented to Ground Targets}
This review surveys dynamic task allocation methods for \acrshort{uav} swarms facing changing ground-target situations. The review classifies dynamic task allocation models into global and local approaches and analyses algorithms including market-based, intelligent optimisation, and clustering strategies. Each method’s advantages and practical performance trade-offs are discussed. The review identifies ongoing challenges such as communication limits and heterogeneous \acrshort{uav} swarms, and suggests future integration of AI and distributed optimisation for adaptive task allocation.

\subsection{UAV swarm communication and control architectures: a review}
This paper reviews UAV swarm communication and control architectures by comparing centralised (ground-based) and \gls{adhoc} (\acrshort{fanet}) models. The paper discusses autonomy levels, swarm coordination algorithms, and the potential of 5G networks to improve UAV-to-UAV communication. The paper highlights the need for hybrid architectures that combine distributed decision-making with reliable infrastructure support to overcome range, latency, and scalability issues.

\subsection{Summary}
In summary, existing research demonstrates that UAV swarms can coordinate search patterns, maintain basic connectivity in flying ad-hoc networks, and perform dynamic task allocation under various assumptions. However, most SAR-oriented swarm solutions either assume fully functional agents or handle failures by removing compromised units without explicitly restructuring the mission. Consensus-oriented works focus on agreement but do not couple consensus with health-aware task reallocation, while task-allocation studies often neglect detailed models of agent health and fault propagation. None of the reviewed works provides an integrated, fail-operational replanning protocol that combines distributed consensus on degraded health events, dynamic redistribution of search sectors, and assignment of secondary roles to compromised agents in a SAR context. Addressing this gap is the central focus of the IRDS project.