\section{Project Context \& Stakeholders}

% Main Roles: CE, RM, SM, VVM, QCM
% Support Roles: 
% - Requirements Engineer
% - System Architect
% - Software/Hardware Architect
% - Software/Hardware Designer
% - Developer / System Developer
% - Tester (Software/Hardware Tester)
% - Quality Engineer
% - Configuration Engineer
% - Safety Engineer

\subsection{Team Members}
\Cref{tab:team-members} lists people included in the project team.
\renewcommand{\arraystretch}{1.2}
\begin{table}[H]
    \centering
    \begin{tabular}{|l|l|l|}
        \hline
        \textbf{Name} & \textbf{Main Role} & \textbf{Support Role} \\
        \hline
        Andrea Haglund & \acrfull{ce} & System Architect \& Designer \\
        \hline
        Claire Namatovu & \acrfull{rm} & Requirements Engineer \\
        \hline
        Emily Zainali & \acrfull{vvm} & Tester \\
        \hline
        Esaias Målqvist & \acrfull{sm} & Safety Engineer \\
        \hline
        Yonatan Michael Beyene & \acrfull{qcm} & Quality Engineer \& Configuration Engineer \\
        \hline
    \end{tabular}
    \caption{Project team members.}
    \label{tab:team-members}
\end{table}
\renewcommand{\arraystretch}{1.0}

\subsection{Stakeholders}
\Cref{tab:stakeholders} lists stakeholders related to this project.
\begin{table}[H]
    \centering
    \begin{tabular}{|l|l|l|l|}
        \hline
        \textbf{Name} & \textbf{Role} & \textbf{Responsibilities} & \textbf{Contact Info} \\
        \hline
        Luiz Giacomossi & \makecell{Project Owner} & \makecell{Defines project objectives,\\ approves scope and decisions,\\ ensures expected value} & luiz.giacomossi@mdu.se \\
        \hline
        Luciana Provenzano & \makecell{Course Coordinator} & \makecell{Ensures compliance with course\\ framework, provides academic\\ oversight, evaluates results} & luciana.provenzano@mdu.se \\
        \hline
        Julieth Castellanos & Project Advisor & \makecell{Provides guidance and feedback,\\ supports theory application,\\ ensures academic standards} & julieth.castellanos@mdu.se \\
        \hline
    \end{tabular}
    \caption{Stakeholders.}
    \label{tab:stakeholders}
\end{table}

\subsection{SWOT Analysis}
\textbf{Strengths}
\begin{itemize}
    \item All team members are deeply committed to this project.
    \item The team collectively has a good level of competency and knowledge.
    \item Team members have previous experiences that are advantageous for this project.
\end{itemize}
\textbf{Opportunities}
\begin{itemize}
    %\item \acrfull{uav} swarm \acrfull{sar} is a not deeply explored area.
    %\item Opportunity to broaden team members’ professional networks.
    \item \textit{All team members are deeply committed to this project.}
    \begin{itemize}
        \item[$\circ$] Higher likelihood of meeting project goals due to strong motivation.
        \item[$\circ$] Ability to take on ambitious tasks or project scopes because of strong teamwork.
        \item[$\circ$] Faster problem-solving as members are willing to put in extra effort.
        \item[$\circ$] Increased resilience during setbacks, enabling smoother project progress.
    \end{itemize}
    \item \textit{The team collectively has a good level of competency and knowledge.}
    \begin{itemize}
        \item[$\circ$] Ability to develop high-quality outputs or innovative solutions.
        \item[$\circ$] Reduced need for external assistance, saving time and resources.
        \item[$\circ$] Capability to explore advanced techniques or new technologies in the project.
        \item[$\circ$] Faster learning curves and adaptability to new requirements.
    \end{itemize}
    \item \textit{Team members have previous experiences that are advantageous for this project.}
    \begin{itemize}
        \item[$\circ$] Ability to avoid past mistakes and apply proven strategies.
        \item[$\circ$] Faster execution due to familiarity with similar tasks or processes.
        \item[$\circ$] Improved decision-making based on real-world insights.
        \item[$\circ$] Opportunity to leverage prior networks, tools, or methodologies.
        \item[$\circ$] Greater confidence in handling complex or unexpected challenges.
    \end{itemize}
\end{itemize}
\textbf{Weaknesses}
\begin{itemize}
    \item Lack of experience in project planning and management.
    \item Time constraint.
    \item Lack of hardware specification.
\end{itemize}
\textbf{Threats}
\begin{itemize}
    %\item Sudden illness among team members.
    \item \textit{Lack of experience in project planning and management.}
    \begin{itemize}
        \item[$\circ$] Project delays due to poor task estimation and scheduling.
        \item[$\circ$] Increased risk of project failure due to inadequate oversight or coordination.
        \item[$\circ$] Communication breakdowns within the team.
        \item[$\circ$] Difficulty adapting to unexpected issues, causing further setbacks.
    \end{itemize}
    \item \textit{Time constraint.}
    \begin{itemize}
        \item[$\circ$] Rushed deliverables that may compromise quality and safety.
        \item[$\circ$] Inability to complete all required features or tasks.
        \item[$\circ$] Higher stress levels leading to mistakes or burnout.
        \item[$\circ$] Reduced ability to test, validate, or refine the project.
        \item[$\circ$] Missed deadlines impacting stakeholders’ trust or project approval.
    \end{itemize}
    \item \textit{Lack of hardware specification.}
    \begin{itemize}
        \item[$\circ$] Incompatibility between project requirements and hardware.
    \end{itemize}
\end{itemize}

\subsection{Dependencies \& External Factors}
%- External factors (regulations, course guidelines)
%- Dependencies on other systems or resources
%- High-level risks or assumptions
This project is influenced by several external factors and dependencies that may affect scheduling, risk, or deliverables:
\begin{itemize}
    \item Deadlines, Working Days and Hours
    \begin{itemize}
        \item[$\circ$] This project must align with submission deadlines and seminar dates.
        \item[$\circ$] Delays in intermediate activities (e.g., requirements validation and simulation results) could affect final delivery.
        \item[$\circ$] The project schedule is based on regular weekdays and standard working hours, and work is assumed to progress only during these periods, excluding weekends and public holidays.
        \item[$\circ$] Any deviations (e.g., reduced team availability) could delay milestone achievement.
    \end{itemize}
    \item Simulation Software Availability
    \begin{itemize}
        \item[$\circ$] This project depends on access to a specific \acrshort{uav} simulation software (gym-pybullets-drones) that will be modified according to project needs.
        \item[$\circ$] If modification of the software is delayed, testing may be delayed or scaled back to adhere to delivery dates.
    \end{itemize}
    \item Reference to \acrshort{uav} Specifications
    \begin{itemize}
        \item[$\circ$] Requirements and performance constraints are based on \href{https://harrisaerial.com/carrier-drones/carrier-h6hl/}{Harris Aerial’s Carrier H6HL} specifications.
        \item[$\circ$] If those specifications are updated or incomplete, assumptions may have to be revised.
    \end{itemize}
    \item External Regulations \& Standards
    \begin{itemize}
        \item[$\circ$] Although no physical \acrshort{uav} will be built, simulations must still follow standard \acrshort{uav} safety and performance guidelines for validity.
        \item[$\circ$] Changes in regulatory requirements may impact this project.
    \end{itemize}
\end{itemize}
Several of these dependencies and external factors present potential risks to this project’s timeline and quality. For example, delays in modification of the simulation software could directly impact milestone achievement. These risks, along with mitigation strategies, are further detailed in \nameref{section:risk}.